### 2023-06-05 Code/BatchEffect.Rmd
Added simulation code for 2 random effects under BatchEffect.Rmd

---
title: "Random_Effects"
author: "Owen Fiore"
date: "2023-05-31"
output: pdf_document
---

```{r}
library(lme4)
two_effect <- glmer(RxnTime ~ (1 | Year) + (1 | Batch), data = men_pooled, family = Gamma(link = "log"))
one_effect <- glmer(RxnTime ~ (1 | Year) , data = men_pooled, family = Gamma(link = "log"))
no_effect <- glm(RxnTime ~ 1, data = men_pooled, family = Gamma(link = "log"))
batch <- glmer(RxnTime ~ (1 | Batch) , data = men_pooled, family = Gamma(link = "log"))
anova(batch, two_effect)
print("------------------")
anova(one_effect, two_effect)
print("--------------")
par(mfrow=c(1,1))
qqnorm(resid(two_effect))
par(mfrow=c(2,1))
qqnorm(ranef(two_effect)$Year[,1])
qqnorm(ranef(two_effect)$Batch[,1])
qqnorm(ranef(batch)$Batch[,1])
qqnorm(resid(one_effect), main = "Without Batch Effect")
#anova(one_effect, two_effect)
```

```{r}
nested <- glmer(RxnTime ~ (1|Year) + (1|Batch:Year), data=men_pooled, family = Gamma(link="log"))
anova(nested, one_effect)
batch_year <- glmer(RxnTime ~ (1|Batch:Year), data=men_pooled, family = Gamma(link="log"))
anova(batch_year, nested)
#ranef(nested)
```

```{r}
simulation <- function(model, base, RxnTime) {
  simfit <- function(model, n = 10000000) {
    vc <- as.data.frame(VarCorr(model))
    z1 <- rnorm(n, sd = vc$sdcor[1])  # random effects for Year
    z2 <- rnorm(n, sd = vc$sdcor[2])  # random effects for Batch
    mu <- exp(fixef(model) + z1 + z2)
    shape <- 1 / vc$vcov[2]
    x <- rgamma(n, shape = shape, scale = mu / shape)
    x
  }

  sbase <- summary(base)
  x <- simfit(model)
  print(mean(x < RxnTime))
  print(quantile(x, prob = c(0.00001, 0.0001, 0.0005, 0.001)))
  ## compare with and without random effect
  plot(density(x))  # with random effects
  curve(dgamma(x, shape = 1 / sbase$dispersion, 
               scale = exp(coef(base)) * sbase$dispersion), 
        0.08, 0.22, add = TRUE, col = "blue")  # fixed effect only
}
simulation(old_pooled_year_gamma_log, men_pooled_one_gamma_log, 0.1)
```
```{}
All probabilities are for RxnTime=0.1 seconds
old pooled: 0.0283374 vs 0.0057613
men_pooled: 0.0313561 vs 0.0068649
old finals: 0.0056892 vs 0.0004389
men_finals: 0.0083227 vs 0.000856
The times on the left are for the 2 batch, times on the right are for the 1 batch
```