### 2023-09-06 Code/GLMM.Rmd
Started to remove finals variables and renamed Batch to Heat

---
title: "GLMM"
author: "Owen Fiore"
date: "2023-05-31"
output: pdf_document
---

Professors Yan and Schifano: to see the probabilities of the three different models
please look at the results from code chunks 7, 12, 13.  You will need to run
code from other chunks to generate the models and for the simulation code, but
these three will produce probabilities of a reaction time less than 0.1 seconds.

The code below loads in the dataset from the csv file.
```{r}
#1
times <- read.csv("../Data/Thesis2023.csv")
colnames(times) <- c("Year", "Type", "Time", "RxnTime", "Gender", "Heat")
men_pooled <- subset(times, (Type == "F" | Type == "S") & 
                       Gender == "M" & Time != 'DQ')
men_old_pooled <- subset(times, (Type == "F" | Type == "S") & 
                           Gender == "M" & Time != 'DQ' & Year != 2022)
```

This code below creates the base models without any random effects.  These are used
as a baseline to compare the year effect models to in the simulation code below.
```{r}
#2
library(lme4)
men_old_pooled_one_gamma_log <- glm(RxnTime ~ 1, data = men_old_pooled, 
                                    family = Gamma(link = "log"))
men_pooled_one_gamma_log <- glm(RxnTime ~ 1, data = men_pooled, 
                                family = Gamma(link = "log"))
```

This code creates the year effect only models.
```{r}
#3
library(lme4)
men_old_pooled_year_gamma_log <- glmer(RxnTime ~ (1|Year), data = men_old_pooled, 
                                       family = Gamma(link = "log"))
men_pooled_year_gamma_log <- glmer(RxnTime ~ (1|Year), data = men_pooled, 
                                   family = Gamma(link = "log"))
```

```{r}
summary(men_pooled_year_gamma_log)
summary(men_old_pooled_year_gamma_log)
```

This creates the heat effect only models.
```{r}
#4
library(lme4)
men_old_pooled_heat_gamma_log <-glmer(RxnTime ~ (1 | Heat), data = men_old_pooled, 
                             family = Gamma(link = "log"))
men_pooled_heat_gamma_log <-glmer(RxnTime ~ (1 | Heat), data = men_pooled, 
                         family = Gamma(link = "log"))
```

```{r}
summary(men_pooled_heat_gamma_log)
```

This creates the models with a year effect and heat effect.
```{r}
#5
library(lme4)
men_old_finals_two_effect_gamma_log <- glmer(RxnTime ~ (1 | Year) + (1 | heat), 
                                   data = men_old_finals, family = Gamma(link = "log"))
men_finals_two_effect_gamma_log<- glmer(RxnTime ~ (1 | Year) + (1 | heat), 
                              data = men_finals, family = Gamma(link = "log"))
men_old_pooled_two_effect_gamma_log <- glmer(RxnTime ~ (1 | Year) + (1 | heat), 
                                   data = men_old_pooled, family = Gamma(link = "log"))
men_pooled_two_effect_gamma_log <- glmer(RxnTime ~ (1 | Year) + (1 | heat), 
                               data = men_pooled, family = Gamma(link = "log"))
```
```{r}
summary(men_pooled_two_effect_gamma_log)
summary(men_old_pooled_two_effect_gamma_log)
```

```{r}
anova(men_pooled_heat_gamma_log, men_pooled_two_effect_gamma_log)
```
This anova test determined that the two effect model is an improvement over the
heat effect model. The test statistic is 0.01059 which is significant at alpha 
= 0.05 but not at alpha = 0.01.

```{r}
anova(men_pooled_year_gamma_log, men_pooled_two_effect_gamma_log)
```
This anova test determined that the two effect model is a significant improvement
over the year effect model.

This code simulates the probability of observing a reaction time less than 0.1
for the various data sets.  The function takes in two models and a reaction time,
it is important that the two models that are compared are from the same data set;
we want to compare old_finals_two_effect and old_finals_one_gamma_log.
```{r}
#6
set.seed(1)
library(lme4)
sim_two_effect_time <- function(model, base, RxnTime) {
  simfit <- function(model, n = 10000000) {
    vc <- as.data.frame(VarCorr(model))
    sd_year <- vc$sdcor[1]
    sd_year_heat <- vc$sdcor[2]

    z1 <- rnorm(n, sd = sd_year)
    z2 <- rnorm(n, sd = sd_year_heat)

    shape <- 1 / sigma(model)^2
    mu <- exp(fixef(model) + z1 + z2)

    x <- rgamma(n, shape = shape, scale = mu / shape)
    x
  }

  sbase <- summary(base)
  x <- simfit(model)
  print(mean(x < RxnTime))
  #print(quantile(x, prob = c(0.00001, 0.0001, 0.0005, 0.001)))
  ## compare with and without random effect
  plot(density(x))  # with random effects
  curve(dgamma(x, shape = 1 / sbase$dispersion, 
               scale = exp(coef(base)) * sbase$dispersion), 
        0.08, 0.25, add = TRUE, col = "blue")  # fixed effect only
}
```

This next chunk of code runs the simulation code for each dataset comparing the
two effect model to the base model
```{r}
#7
set.seed(1)
sim_two_effect_time(men_old_pooled_two_effect_gamma_log, men_old_pooled_one_gamma_log, 0.1)
sim_two_effect_time(men_pooled_two_effect_gamma_log, men_pooled_one_gamma_log, 0.1)
```

```{r}
set.seed(1)
sim_two_effect_time(men_old_pooled_two_effect_gamma_log, men_old_pooled_one_gamma_log, 0.09)
sim_two_effect_time(men_pooled_two_effect_gamma_log, men_pooled_one_gamma_log, 0.09)
```

```{r}
set.seed(1)
sim_two_effect_time(men_old_pooled_two_effect_gamma_log, men_old_pooled_one_gamma_log, 0.08)
sim_two_effect_time(men_pooled_two_effect_gamma_log, men_pooled_one_gamma_log, 0.08)
```

sim_two_effect_prob is a slight modification of the earlier function that instead
takes as input a probability rather than a time.  The output is a reaction time 
barrier that allows a reaction time below the barrier one in every 1/(input probability)
times. Decreasing the probability will impose a stricter barrier as it is less
probable to observe a reaction time of 0.08 rather than 0.09.  Below are the
recommended reaction time barriers to disqualify on average 1 in every 1,000 starts.
```{r}
#8
set.seed(1)
library(lme4)
sim_two_effect_prob <- function(model, base, prob) {
  simfit <- function(model, n = 10000000) {
    vc <- as.data.frame(VarCorr(model))
    sd_year <- vc$sdcor[1]
    sd_year_heat <- vc$sdcor[2]

    z1 <- rnorm(n, sd = sd_year)
    z2 <- rnorm(n, sd = sd_year_heat)

    shape <- 1 / sigma(model)^2
    mu <- exp(fixef(model) + z1 + z2)

    x <- rgamma(n, shape = shape, scale = mu / shape)
    x
  }

  sbase <- summary(base)
  x <- simfit(model)
  print(quantile(x, probs = prob))
}
```

This runs the sim_two_effect_prob function at a probability of 1 in every 1,000
sprinters.
```{r}
#9
set.seed(1)
sim_two_effect_prob(men_old_pooled_two_effect_gamma_log, men_old_pooled_one_gamma_log, 0.001)
sim_two_effect_prob(men_pooled_two_effect_gamma_log, men_pooled_one_gamma_log, 0.001)
```

The code below repeats the previous section but for an average of 1 in every
10,000 sprinters.
```{r}
#10
set.seed(1)
sim_two_effect_prob(men_old_pooled_two_effect_gamma_log, men_old_pooled_one_gamma_log, 0.0001)
sim_two_effect_prob(men_pooled_two_effect_gamma_log, men_pooled_one_gamma_log, 0.0001)
```

This next code section replicates the code above but the various simulation
functions are meant for one random effect instead of two.
```{r}
#11
library(lme4)
set.seed(1)
sim_one_effect_time <- function(model, base, RxnTime){
    simfit <- function(model, n = 10000000) {
    vc <- as.data.frame(VarCorr(model))
    ## mu <- exp(coef(model)$Year[,]) ## using fitted random effects only
    z <- rnorm(n, sd = vc$sdcor[1]) #$ using generated random effects
    mu <- exp(fixef(model) + z)
    shape <- 1 / vc$vcov[2]
    x <- rgamma(n, shape = shape, scale = mu / shape)
    x
    }
  sbase <- summary(base)
  #print(sbase)
  x <- simfit(model)
  print(mean(x < RxnTime))
  ## compare with and without random effect
  plot(density(x)) # with random effect
  curve(dgamma(x, shape = 1 / sbase$dispersion, 
               scale = exp(coef(base)) * sbase$dispersion), 
        0.08, 0.25, add = TRUE, col = "blue") # fixed effect only
}
```

```{r}
#12
sim_one_effect_time(men_old_finals_year_gamma_log, men_old_finals_one_gamma_log, 0.1)
sim_one_effect_time(men_finals_year_gamma_log, men_finals_one_gamma_log, 0.1)
sim_one_effect_time(men_old_pooled_year_gamma_log, men_old_pooled_one_gamma_log, 0.1)
sim_one_effect_time(men_pooled_year_gamma_log, men_pooled_one_gamma_log, 0.1)
```

```{r}
#13
sim_one_effect_time(men_old_finals_heat_gamma_log, men_old_finals_one_gamma_log, 0.1)
sim_one_effect_time(men_finals_heat_gamma_log, men_finals_one_gamma_log, 0.1)
sim_one_effect_time(men_old_pooled_heat_gamma_log, men_old_pooled_one_gamma_log, 0.1)
sim_one_effect_time(men_pooled_heat_gamma_log, men_pooled_one_gamma_log, 0.1)
```

```{r}
#14
library(lme4)
set.seed(1)
sim_one_effect_prob <- function(model, base, prob) {
  simfit <- function(model, n = 10000000) {
    vc <- as.data.frame(VarCorr(model))
    z <- rnorm(n, sd = vc$sdcor[1])
    mu <- exp(fixef(model) + z)
    shape <- 1 / vc$vcov[2]
    x <- rgamma(n, shape = shape, scale = mu / shape)
    x
  }

  sbase <- summary(base)
  x <- simfit(model)
  print(quantile(x, probs = prob))
}
```

```{r}
#15
sim_one_effect_prob(men_old_finals_year_gamma_log, men_old_finals_one_gamma_log, 0.001)
sim_one_effect_prob(men_finals_year_gamma_log, men_finals_one_gamma_log, 0.001)
sim_one_effect_prob(men_old_pooled_year_gamma_log, men_old_pooled_one_gamma_log, 0.001)
sim_one_effect_prob(men_pooled_year_gamma_log, men_pooled_one_gamma_log, 0.001)
```

```{r}
#16
sim_one_effect_prob(men_old_finals_year_gamma_log, men_old_finals_one_gamma_log, 0.0001)
sim_one_effect_prob(men_finals_year_gamma_log, men_finals_one_gamma_log, 0.0001)
sim_one_effect_prob(men_old_pooled_year_gamma_log, men_old_pooled_one_gamma_log, 0.0001)
sim_one_effect_prob(men_pooled_year_gamma_log, men_pooled_one_gamma_log, 0.0001)
```