### 2023-09-13 Code/GLMM.Rmd
Renamed variables

---
title: "GLMM"
author: "Owen Fiore"
date: "2023-05-31"
output:
  html_document:
    df_print: paged
---

To see the probabilities associated with a reaction time below 0.1 seconds for
the two effect model, run code chunks 1, 2, 5, 6, 7.

The code below loads in the data sets from the csv file.
```{r}
#1
times <- read.csv("../Data/Thesis2023.csv")
colnames(times) <- c("Venue", "Type", "Time", "RxnTime", "Gender", "Heat")
with2022 <- subset(times, (Type == "F" | Type == "S") & 
                       Gender == "M" & Time != 'DQ')
without2022 <- subset(times, (Type == "F" | Type == "S") & 
                           Gender == "M" & Time != 'DQ' & Venue != 2022)
```

This code below creates the base models without any random effects.  These are used
as a baseline to compare the venue effect models to in the simulation code below.
```{r}
#2
library(lme4)
without2022_one_gamma_log <- glm(RxnTime ~ 1, data = without2022, 
                                    family = Gamma(link = "log"))
with2022_one_gamma_log <- glm(RxnTime ~ 1, data = with2022, 
                                family = Gamma(link = "log"))
```

This code creates the venue effect only models.
```{r}
#3
library(lme4)
without2022_venue_gamma_log <- glmer(RxnTime ~ (1|Venue), data = without2022, 
                                       family = Gamma(link = "log"))
with2022_venue_gamma_log <- glmer(RxnTime ~ (1|Venue), data = with2022, 
                                   family = Gamma(link = "log"))
```

```{r}
summary(with2022_venue_gamma_log)
summary(without2022_venue_gamma_log)
```

This creates the heat effect only models.
```{r}
#4
library(lme4)
without2022_heat_gamma_log <-glmer(RxnTime ~ (1 | Heat), data = without2022, 
                             family = Gamma(link = "log"))
with2022_heat_gamma_log <-glmer(RxnTime ~ (1 | Heat), data = with2022, 
                         family = Gamma(link = "log"))
```

```{r}
summary(with2022_heat_gamma_log)
```

This creates the models with a venue effect and heat effect.
```{r}
#5
library(lme4)
without2022_two_effect_gamma_log <- glmer(RxnTime ~ (1 | Venue) + (1 | Heat), 
                                   data = without2022, family = Gamma(link = "log"))
with2022_two_effect_gamma_log <- glmer(RxnTime ~ (1 | Venue) + (1 | Heat), 
                               data = with2022, family = Gamma(link = "log"))
```

```{r}
summary(with2022_two_effect_gamma_log)
summary(without2022_two_effect_gamma_log)
```

```{r}
anova(with2022_heat_gamma_log, with2022_two_effect_gamma_log)
```
This anova test determined that the two effect model is an improvement over the
heat effect model. The test statistic is 0.01059 which is significant at alpha 
= 0.05 but not at alpha = 0.01.

```{r}
anova(with2022_venue_gamma_log, with2022_two_effect_gamma_log)
```
This anova test determined that the two effect model is a significant improvement
over the Venue effect model.

This code simulates the probability of observing a reaction time less than 0.1
for the various data sets.  The function takes in two models and a reaction time,
it is important that the two models that are compared are from the same data set;
we want to compare old_finals_two_effect and old_finals_one_gamma_log.
```{r}
#6
set.seed(1)
library(lme4)
sim_two_effect_time <- function(model, base, RxnTime) {
  simfit <- function(model, n = 10000000) {
    vc <- as.data.frame(VarCorr(model))
    sd_Venue <- vc$sdcor[1]
    sd_Venue_heat <- vc$sdcor[2]

    z1 <- rnorm(n, sd = sd_Venue)
    z2 <- rnorm(n, sd = sd_Venue_heat)

    shape <- 1 / sigma(model)^2
    mu <- exp(fixef(model) + z1 + z2)

    x <- rgamma(n, shape = shape, scale = mu / shape)
    x
  }

  sbase <- summary(base)
  x <- simfit(model)
  print(mean(x < RxnTime))
  #print(quantile(x, prob = c(0.00001, 0.0001, 0.0005, 0.001)))
  ## compare with and without random effect
  plot(density(x))  # with random effects
  curve(dgamma(x, shape = 1 / sbase$dispersion, 
               scale = exp(coef(base)) * sbase$dispersion), 
        0.08, 0.25, add = TRUE, col = "blue")  # fixed effect only
}
```

This next chunk of code runs the simulation code for each dataset comparing the
two effect model to the base model
```{r}
#7
set.seed(1)
sim_two_effect_time(without2022_two_effect_gamma_log, without2022_one_gamma_log, 0.1)
sim_two_effect_time(with2022_two_effect_gamma_log, with2022_one_gamma_log, 0.1)
```

```{r}
set.seed(1)
sim_two_effect_time(without2022_two_effect_gamma_log, without2022_one_gamma_log, 0.09)
sim_two_effect_time(with2022_two_effect_gamma_log, with2022_one_gamma_log, 0.09)
```

```{r}
set.seed(1)
sim_two_effect_time(without2022_two_effect_gamma_log, without2022_one_gamma_log, 0.08)
sim_two_effect_time(with2022_two_effect_gamma_log, with2022_one_gamma_log, 0.08)
```

sim_two_effect_prob is a slight modification of the earlier function that instead
takes as input a probability rather than a time.  The output is a reaction time 
barrier that allows a reaction time below the barrier one in every 1/(input probability)
times. Decreasing the probability will impose a stricter barrier as it is less
probable to observe a reaction time of 0.08 rather than 0.09.  Below are the
recommended reaction time barriers to disqualify on average 1 in every 1,000 starts.
```{r}
#8
set.seed(1)
library(lme4)
sim_two_effect_prob <- function(model, base, prob) {
  simfit <- function(model, n = 10000000) {
    vc <- as.data.frame(VarCorr(model))
    sd_Venue <- vc$sdcor[1]
    sd_Venue_heat <- vc$sdcor[2]

    z1 <- rnorm(n, sd = sd_Venue)
    z2 <- rnorm(n, sd = sd_Venue_heat)

    shape <- 1 / sigma(model)^2
    mu <- exp(fixef(model) + z1 + z2)

    x <- rgamma(n, shape = shape, scale = mu / shape)
    x
  }

  sbase <- summary(base)
  x <- simfit(model)
  print(quantile(x, probs = prob))
}
```

This runs the sim_two_effect_prob function at a probability of 1 in every 1,000
sprinters.
```{r}
#9
set.seed(1)
sim_two_effect_prob(without2022_two_effect_gamma_log, without2022_one_gamma_log, 0.001)
sim_two_effect_prob(with2022_two_effect_gamma_log, with2022_one_gamma_log, 0.001)
```

The code below repeats the previous section but for an average of 1 in every
10,000 sprinters.
```{r}
#10
set.seed(1)
sim_two_effect_prob(without2022_two_effect_gamma_log, without2022_one_gamma_log, 0.0001)
sim_two_effect_prob(with2022_two_effect_gamma_log, with2022_one_gamma_log, 0.0001)
```

This next code section replicates the code above but the various simulation
functions are meant for one random effect instead of two.
```{r}
#11
library(lme4)
set.seed(1)
sim_one_effect_time <- function(model, base, RxnTime){
    simfit <- function(model, n = 10000000) {
    vc <- as.data.frame(VarCorr(model))
    ## mu <- exp(coef(model)$Venue[,]) ## using fitted random effects only
    z <- rnorm(n, sd = vc$sdcor[1]) #$ using generated random effects
    mu <- exp(fixef(model) + z)
    shape <- 1 / vc$vcov[2]
    x <- rgamma(n, shape = shape, scale = mu / shape)
    x
    }
  sbase <- summary(base)
  #print(sbase)
  x <- simfit(model)
  print(mean(x < RxnTime))
  ## compare with and without random effect
  plot(density(x)) # with random effect
  curve(dgamma(x, shape = 1 / sbase$dispersion, 
               scale = exp(coef(base)) * sbase$dispersion), 
        0.08, 0.25, add = TRUE, col = "blue") # fixed effect only
}
```

```{r}
#12
set.seed(1)
sim_one_effect_time(without2022_venue_gamma_log, without2022_one_gamma_log, 0.1)
sim_one_effect_time(with2022_venue_gamma_log, with2022_one_gamma_log, 0.1)
```

```{r}
#13
set.seed(1)
sim_one_effect_time(without2022_heat_gamma_log, without2022_one_gamma_log, 0.1)
sim_one_effect_time(with2022_heat_gamma_log, with2022_one_gamma_log, 0.1)
```

```{r}
#14
library(lme4)
set.seed(1)
sim_one_effect_prob <- function(model, base, prob) {
  simfit <- function(model, n = 10000000) {
    vc <- as.data.frame(VarCorr(model))
    z <- rnorm(n, sd = vc$sdcor[1])
    mu <- exp(fixef(model) + z)
    shape <- 1 / vc$vcov[2]
    x <- rgamma(n, shape = shape, scale = mu / shape)
    x
  }

  sbase <- summary(base)
  x <- simfit(model)
  print(quantile(x, probs = prob))
}
```

```{r}
#15
sim_one_effect_prob(without2022_venue_gamma_log, without2022_one_gamma_log, 0.001)
sim_one_effect_prob(with2022_venue_gamma_log, with2022_one_gamma_log, 0.001)
```

```{r}
#16
sim_one_effect_prob(without2022_venue_gamma_log, without2022_one_gamma_log, 0.0001)
sim_one_effect_prob(with2022_venue_gamma_log, with2022_one_gamma_log, 0.0001)
```