### 2023-10-18 Archive/Models.Rmd
Contains many old models that were at one time considered but not necessary anymore

---
title: "Models"
author: "Owen Fiore"
date: "2023-02-07"
output: pdf_document
---

```{r}
times <- read.csv("../Data/ThesisData.csv")
colnames(times) <- c("Year", "Type", "Time", "RxnTime", "Gender", "Batch")
```

```{r}
men_finals <- subset(times, Type == "F" & Gender == "M" & Time != 'DQ') # & Year < 2022)
men_pooled <- subset(times, (Type == "F" | Type == "S") & Gender == "M" & Time != 'DQ')
```

```{r}
men_old_finals <- subset(times, Type == "F" & Gender == "M" & Time != 'DQ' & Year < 2022)
men_old_pooled <- subset(times, (Type == "F" | Type == "S") & Gender == "M" & Time != 'DQ' & Year < 2022)
```

```{r}
library(lme4)
men_finals_one <- glm(RxnTime ~ 1, data = men_finals)
men_finals_year <- lmer(RxnTime ~ (1|Year), data = men_finals)
men_finals_one_gamma <- glm(RxnTime ~ 1, data = men_finals, family = Gamma)
men_finals_year_gamma <- glmer(RxnTime ~ (1|Year), data = men_finals, family = Gamma)
men_finals_one_gamma_log <- glm(RxnTime ~ 1, data = men_finals, family = Gamma(link = "log"))
men_finals_year_gamma_log <- glmer(RxnTime ~ (1|Year), data = men_finals, family = Gamma(link = "log"))

paste("Linear without year effect has AIC of", AIC(men_finals_one), "and log-lik of", logLik(men_finals_one))
paste("Linear with year effect has AIC of", AIC(men_finals_year), "and log-lik of", logLik(men_finals_year))
paste("Gamma without year effect has AIC of", AIC(men_finals_one_gamma), "and log-lik of", logLik(men_finals_one_gamma))
paste("Gamma with year effect has AIC of", AIC(men_finals_year_gamma), "and log-lik of", logLik(men_finals_year_gamma))
paste("Gamma and log without year effect has AIC of", AIC(men_finals_one_gamma_log), "and log-lik of", logLik(men_finals_one_gamma_log))
paste("Gamma and log with year effect has AIC of", AIC(men_finals_year_gamma_log), "and log-lik of", logLik(men_finals_year_gamma_log))
```

The best selected model is men_finals_year_gamma as that has the smallest AIC: -491 and largest log-lik of 248.5

```{r}
library(lme4)
men_old_finals_one <- glm(RxnTime ~ 1, data = men_old_finals)
men_old_finals_year <- lmer(RxnTime ~ (1|Year), data = men_old_finals)
men_old_finals_one_gamma <- glm(RxnTime ~ 1, data = men_old_finals, family = Gamma)
men_old_finals_year_gamma <- glmer(RxnTime ~ (1|Year), data = men_old_finals, family = Gamma)
men_old_finals_one_gamma_log <- glm(RxnTime ~ 1, data = men_old_finals, family = Gamma(link = "log"))
men_old_finals_year_gamma_log <- glmer(RxnTime ~ (1|Year), data = men_old_finals, family = Gamma(link = "log"))

paste("Linear without year effect has AIC of", AIC(men_old_finals_one), "and log-lik of", logLik(men_old_finals_one))
paste("Linear with year effect has AIC of", AIC(men_old_finals_year), "and log-lik of", logLik(men_old_finals_year))
paste("Gamma without year effect has AIC of", AIC(men_old_finals_one_gamma), "and log-lik of", logLik(men_old_finals_one_gamma))
paste("Gamma with year effect has AIC of", AIC(men_old_finals_year_gamma), "and log-lik of", logLik(men_old_finals_year_gamma))
paste("Gamma and log without year effect has AIC of", AIC(men_old_finals_one_gamma_log), "and log-lik of", logLik(men_old_finals_one_gamma_log))
paste("Gamma and log with year effect has AIC of", AIC(men_old_finals_year_gamma_log), "and log-lik of", logLik(men_old_finals_year_gamma_log))
```

The best selected model is men_old_finals_year_gamma as that has the smallest AIC: -465 and largest log-lik of 235.5.  The models for the older data does not seem to fit the data as good as for the full data set.

```{r}
library(lme4)
men_pooled_one <- glm(RxnTime ~ 1, data = men_pooled)
men_pooled_year <- lmer(RxnTime ~ (1|Year), data = men_pooled)
men_pooled_one_gamma <- glm(RxnTime ~ 1, data = men_pooled, family = Gamma)
men_pooled_year_gamma <- glmer(RxnTime ~ (1|Year), data = men_pooled, family = Gamma)
men_pooled_one_gamma_log <- glm(RxnTime ~ 1, data = men_pooled, family = Gamma(link = "log"))
men_pooled_year_gamma_log <- glmer(RxnTime ~ (1|Year), data = men_pooled, family = Gamma(link = "log"))

paste("Linear without year effect has AIC of", AIC(men_pooled_one), "and log-lik of", logLik(men_pooled_one))
paste("Linear with year effect has AIC of", AIC(men_pooled_year), "and log-lik of", logLik(men_pooled_year))
paste("Gamma without year effect has AIC of", AIC(men_pooled_one_gamma), "and log-lik of", logLik(men_pooled_one_gamma))
paste("Gamma with year effect has AIC of", AIC(men_pooled_year_gamma), "and log-lik of", logLik(men_pooled_year_gamma))
paste("Gamma and log without year effect has AIC of", AIC(men_pooled_one_gamma_log), "and log-lik of", logLik(men_pooled_one_gamma_log))
paste("Gamma and log with year effect has AIC of", AIC(men_pooled_year_gamma_log), "and log-lik of", logLik(men_pooled_year_gamma_log))
```

The best selected model is men_pooled_year_gamma_log as it had the smallest AIC and the largest log-lik

```{r}
library(lme4)
men_old_pooled_one <- glm(RxnTime ~ 1, data = men_old_pooled)
men_old_pooled_year <- lmer(RxnTime ~ (1|Year), data = men_old_pooled)
men_old_pooled_one_gamma <- glm(RxnTime ~ 1, data = men_old_pooled, family = Gamma)
men_old_pooled_year_gamma <- glmer(RxnTime ~ (1|Year), data = men_old_pooled, family = Gamma)
men_old_pooled_one_gamma_log <- glm(RxnTime ~ 1, data = men_old_pooled, family = Gamma(link = "log"))
men_old_pooled_year_gamma_log <- glmer(RxnTime ~ (1|Year), data = men_old_pooled, family = Gamma(link = "log"))

paste("Linear without year effect has AIC of", AIC(men_old_pooled_one), "and log-lik of", logLik(men_old_pooled_one))
paste("Linear with year effect has AIC of", AIC(men_old_pooled_year), "and log-lik of", logLik(men_old_pooled_year))
paste("Gamma without year effect has AIC of", AIC(men_old_pooled_one_gamma), "and log-lik of", logLik(men_old_pooled_one_gamma))
paste("Gamma with year effect has AIC of", AIC(men_old_pooled_year_gamma), "and log-lik of", logLik(men_old_pooled_year_gamma))
paste("Gamma and log without year effect has AIC of", AIC(men_old_pooled_one_gamma_log), "and log-lik of", logLik(men_old_pooled_one_gamma_log))
paste("Gamma and log with year effect has AIC of", AIC(men_old_pooled_year_gamma_log), "and log-lik of", logLik(men_old_pooled_year_gamma_log))
```

The best model is men_old_pooled_year_gamma as that has the smallest AIC: -1687.5 and the largest log-lik: 846.8

#Trying out identity link
```{r}
men_finals_year_gamma_log <- glmer(RxnTime ~ (1|Year), data = men_finals, family = Gamma(link = "log"))
t1 <- glmer(RxnTime ~ (1|Year), data = men_finals, family = Gamma(link = "identity"))
paste("Gamma and log with year effect has AIC of", AIC(men_finals_year_gamma_log), "and log-lik of", logLik(men_finals_year_gamma_log))
paste("Linear without year effect has AIC of", AIC(t1), "and log-lik of", logLik(t1))
```

```{r}
men_old_finals_year_gamma_log <- glmer(RxnTime ~ (1|Year), data = men_old_finals, family = Gamma(link = "log"))
t2 <- glmer(RxnTime ~ (1|Year), data = men_old_finals, family = Gamma(link = "identity"))
paste("Linear without year effect has AIC of", AIC(men_old_finals_year_gamma_log), "and log-lik of", logLik(men_old_finals_year_gamma_log))
paste("Linear without year effect has AIC of", AIC(t2), "and log-lik of", logLik(t2))
```

```{r}
men_pooled_year_gamma_log <- glmer(RxnTime ~ (1|Year), data = men_pooled, family = Gamma(link = "log"))
t3 <- glmer(RxnTime ~ (1|Year), data = men_pooled, family = Gamma(link = "identity"))
paste("Gamma and log with year effect has AIC of", AIC(men_pooled_year_gamma_log), "and log-lik of", logLik(men_pooled_year_gamma_log))
paste("Linear without year effect has AIC of", AIC(t3), "and log-lik of", logLik(t3))
```

```{r}
men_old_pooled_year_gamma_log <- glmer(RxnTime ~ (1|Year), data = men_old_pooled, family = Gamma(link = "log"))
t4 <- glmer(RxnTime ~ (1|Year), data = men_old_pooled, family = Gamma(link = "identity"))
paste("Gamma and log with year effect has AIC of", AIC(men_old_pooled_year_gamma_log), "and log-lik of", logLik(men_old_pooled_year_gamma_log))
paste("Linear without year effect has AIC of", AIC(t4), "and log-lik of", logLik(t4))

```