### 2023-10-26 Code/GLMM.Rmd
Added a lot more text, explanation, re-organized

---
title: "GLMM"
author: "Owen Fiore"
date: "2023-05-31"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

### This is the code file that shows our results. Code chunks have been labeled for convience.

All outputs rely on code chunks 1-5, run these first.
To see the outputs of Table 2 run code chunks 6-8
To see the outputs of Table 2 run code chunks 11-14
To see the outputs of Table 4 run code chunks 15-17

### 1
The code below loads in the data sets from the csv file.
```{r}

times <- read.csv("../Data/GLMMData.csv")
colnames(times) <- c("Venue", "Type", "Time", "RxnTime", "Gender", "Heat")
with2022 <- subset(times, (Type == "F" | Type == "S") & 
                       Gender == "M" & Time != 'DQ')
without2022 <- subset(times, (Type == "F" | Type == "S") & 
                           Gender == "M" & Time != 'DQ' & Venue != 2022)
```

### 2
This code below creates the null models without any random effects.  These are
used as a baseline to compare the venue effect models to in the simulation code
below.
```{r}
library(lme4)
without2022_one_gamma_log <- glm(RxnTime ~ 1, data = without2022, 
                                    family = Gamma(link = "log"))
with2022_one_gamma_log <- glm(RxnTime ~ 1, data = with2022, 
                                family = Gamma(link = "log"))
```

### 3
This code creates the venue effect only models.
```{r}
library(lme4)
without2022_venue_gamma_log <- glmer(RxnTime ~ (1|Venue), data = without2022, 
                                       family = Gamma(link = "log"))
with2022_venue_gamma_log <- glmer(RxnTime ~ (1|Venue), data = with2022, 
                                   family = Gamma(link = "log"))
```

### 5
This creates the heat effect only models.
```{r}
library(lme4)
without2022_heat_gamma_log <-glmer(RxnTime ~ (1 | Heat), data = without2022, 
                             family = Gamma(link = "log"))
with2022_heat_gamma_log <-glmer(RxnTime ~ (1 | Heat), data = with2022, 
                         family = Gamma(link = "log"))
```

### 5
This creates the models with a venue effect and heat effect.
```{r}
library(lme4)
without2022_two_effect_gamma_log <- glmer(RxnTime ~ (1 | Venue) + (1 | Heat), 
                                   data = without2022,
                                   family = Gamma(link = "log"))
with2022_two_effect_gamma_log <- glmer(RxnTime ~ (1 | Venue) + (1 | Heat), 
                               data = with2022,
                               family = Gamma(link = "log"))
```

### 6
This shows a summary of the venue effect only models.
```{r}
summary(with2022_venue_gamma_log)
summary(without2022_venue_gamma_log)
```

### 7
This shows a summary of the heat effect only models.
```{r}
summary(with2022_heat_gamma_log)
summary(without2022_heat_gamma_log)
```

### 8
This shows a summary of the two effect models.
```{r}
summary(with2022_two_effect_gamma_log)
summary(without2022_two_effect_gamma_log)
```

### 9
This test compares the full model with the heat effect only model.
```{r}
#9
anova(with2022_heat_gamma_log, with2022_two_effect_gamma_log)
```
This anova test determined that the two effect model is an improvement over the
heat effect model. The test statistic is 0.008305 which is significant at 
$\alpha = 0.05$.

### 10
This test compares the full model with the venue effect only model.
```{r}
#10
anova(with2022_venue_gamma_log, with2022_two_effect_gamma_log)
```
This anova test determined that the two effect model is a significant
improvement over the Venue effect model.

### 11
This code simulates the probability of observing a reaction time less than 0.1
for the various data sets.  The function takes in two models and a reaction
time, it is important that the two models that are compared are from the same
data set; we want to compare `old_finals_two_effect` and 
`old_finals_one_gamma_log`.
```{r}
#6
set.seed(1)
library(lme4)
sim_two_effect_time <- function(model, base, RxnTime) {
  simfit <- function(model, n = 10000000) {
    vc <- as.data.frame(VarCorr(model))
    sd_Venue <- vc$sdcor[1]
    sd_Venue_heat <- vc$sdcor[2]

    z1 <- rnorm(n, sd = sd_Venue)
    z2 <- rnorm(n, sd = sd_Venue_heat)

    shape <- 1 / sigma(model)^2
    mu <- exp(fixef(model) + z1 + z2)

    x <- rgamma(n, shape = shape, scale = mu / shape)
    x
  }

  sbase <- summary(base)
  x <- simfit(model)
  print(mean(x < RxnTime))
  #print(quantile(x, prob = c(0.00001, 0.0001, 0.0005, 0.001)))
  ## compare with and without random effect
  plot(density(x))  # with random effects
  curve(dgamma(x, shape = 1 / sbase$dispersion, 
               scale = exp(coef(base)) * sbase$dispersion), 
        0.08, 0.25, add = TRUE, col = "blue")  # fixed effect only
}
```

### 12
This next chunk of code runs the simulation code for each dataset comparing the
two effect model to the base model. We run this code at 3 reaction time barriers:
0.08, 0.09, and 0.1 seconds.
```{r}
#7
set.seed(1)
sim_two_effect_time(without2022_two_effect_gamma_log,
                    without2022_one_gamma_log, 0.1)
sim_two_effect_time(with2022_two_effect_gamma_log,
                    with2022_one_gamma_log, 0.1)
```

### 13
```{r}
set.seed(1)
sim_two_effect_time(without2022_two_effect_gamma_log,
                    without2022_one_gamma_log, 0.09)
sim_two_effect_time(with2022_two_effect_gamma_log,
                    with2022_one_gamma_log, 0.09)
```

### 14
```{r}
set.seed(1)
sim_two_effect_time(without2022_two_effect_gamma_log,
                    without2022_one_gamma_log, 0.08)
sim_two_effect_time(with2022_two_effect_gamma_log,
                    with2022_one_gamma_log, 0.08)
```

### 15
`sim_two_effect_prob` is a slight modification of the earlier function that
instead takes as input a probability rather than a time.  The output is a
reaction time barrier that allows a reaction time below the barrier one in every
1/(input probability) times. Decreasing the probability will impose a stricter
barrier as it is less probable to observe a reaction time of 0.08 rather than
0.09.  Below are the recommended reaction time barriers to disqualify on average
1 in every 1,000 starts.
```{r}
#8
set.seed(1)
library(lme4)
sim_two_effect_prob <- function(model, base, prob) {
  simfit <- function(model, n = 10000000) {
    vc <- as.data.frame(VarCorr(model))
    sd_Venue <- vc$sdcor[1]
    sd_Venue_heat <- vc$sdcor[2]

    z1 <- rnorm(n, sd = sd_Venue)
    z2 <- rnorm(n, sd = sd_Venue_heat)

    shape <- 1 / sigma(model)^2
    mu <- exp(fixef(model) + z1 + z2)

    x <- rgamma(n, shape = shape, scale = mu / shape)
    x
  }

  sbase <- summary(base)
  x <- simfit(model)
  print(quantile(x, probs = prob))
}
```

### 16
This runs the `sim_two_effect_prob` function at a probability of 1 in every 
1,000 sprinters.
```{r}
#9
set.seed(1)
sim_two_effect_prob(without2022_two_effect_gamma_log,
                    without2022_one_gamma_log, 0.001)
sim_two_effect_prob(with2022_two_effect_gamma_log,
                    with2022_one_gamma_log, 0.001)
```

### 17
The code below repeats the previous section but for an average of 1 in every
10,000 sprinters.
```{r}
#10
set.seed(1)
sim_two_effect_prob(without2022_two_effect_gamma_log,
                    without2022_one_gamma_log, 0.0001)
sim_two_effect_prob(with2022_two_effect_gamma_log,
                    with2022_one_gamma_log, 0.0001)
```