### 2024-07-09 Code/NewAnalysis.Rmd
Here I test trying to pool dash and hurdles

---
title: "NewAnalysis"
author: "Owen Fiore"
date: "2024-07-03"
output: html_document
---

# New Work
To try and keep everything organized, we are going to use this document to test
anything additional before putting it in `GLMM.Rmd.`  We have several things
to test before our re-submission.

## Can we pool the dash and hurdles results?
We need to see how different our results are using a pooled version of the men's
100 meter dash and men's 110 meter hurdles.  We will also include a random effect
into our model to compensate for this.

```{r}
library(tidyverse)
times <- read.csv("../Data/GLMMData.csv")
colnames(times) <- c("Venue", "Type", "Time", "RxnTime", "Gender", "Heat", "Event")
```

```{r}
dash_and_hurdles <- times |>
  filter(Type == "F" | Type == "S") |>
  filter(Gender == "M") |>
  filter(Time != "DNS") |>
  filter(RxnTime != 0) |>
  filter(Event == "100 Dash" | Event == "110 Hurdles") |>
  mutate(RxnTime = ifelse(RxnTime < 0.1, 0.1, RxnTime))

hurdles <- dash_and_hurdles |>
  filter(Event == "110 Hurdles")
```

We consider the baseline models:
```{r}
library(lme4)
dash_and_hurdles_one_gamma_log <- glm(RxnTime ~ 1, data = dash_and_hurdles, 
                                    family = Gamma(link = "log"))
hurdles_one_gamma_log <- glm(RxnTime ~ 1, data = hurdles, 
                                family = Gamma(link = "log"))
```
To assess whether we we can pool the data, we are going to perform an anova test
to determine whether the most recently added coefficient to the model was
significant.

```{r}
dash_and_hurdles_three_effect_gamma_log <- glmer(RxnTime ~ (1 | Venue) +
                                                   (1 | Heat) + (1 | Event), 
                                   data = dash_and_hurdles,
                                   family = Gamma(link = "log"))
dash_and_hurdles_two_effect_gamma_log <- glmer(RxnTime ~ (1 | Venue) +
                                                   (1 | Heat), 
                                   data = dash_and_hurdles,
                                   family = Gamma(link = "log"))

hurdles_two_effect_gamma_log <- glmer(RxnTime ~ (1 | Venue) + (1 | Heat), 
                               data = hurdles,
                               family = Gamma(link = "log"))
```

```{r}
anova(dash_and_hurdles_two_effect_gamma_log, dash_and_hurdles_three_effect_gamma_log)
```
What we observe is that there is not a significant difference between these
models as the p value is quite large.

Lets compare the two models
```{r}
summary(dash_and_hurdles_three_effect_gamma_log)
summary(hurdles_two_effect_gamma_log)
```
Now we will see how the inclusion of the pooled data will influence the results
```{r}
sim_three_effect_time <- function(model, base, RxnTime) {
  simfit <- function(model, n = 10000000) {
    vc <- as.data.frame(VarCorr(model))
    sd_Venue <- vc$sdcor[1]
    sd_Venue_heat <- vc$sdcor[2]
    sd_event <- vc$sdcor[3]

    z1 <- rnorm(n, sd = sd_Venue)
    z2 <- rnorm(n, sd = sd_Venue_heat)
    z3 <- rnorm(n, sd = sd_event)
    shape <- 1 / sigma(model)^2
    mu <- exp(fixef(model) + z1 + z2 + z3)

    x <- rgamma(n, shape = shape, scale = mu / shape)
    x
  }

  sbase <- summary(base)
  x <- simfit(model)
  print(mean(x < RxnTime))
  #print(quantile(x, prob = c(0.00001, 0.0001, 0.0005, 0.001)))
  ## compare with and without random effect
  plot(density(x))  # with random effects
  curve(dgamma(x, shape = 1 / sbase$dispersion, 
               scale = exp(coef(base)) * sbase$dispersion), 
        0.08, 0.25, add = TRUE, col = "blue")
}
```

Now we test
```{r}
set.seed(1)
sim_three_effect_time(dash_and_hurdles_three_effect_gamma_log, dash_and_hurdles_one_gamma_log, 0.1)
```