### 2024-10-15 Code/RankComparison.Rmd
Re-organized the code, but still need to run the permutation tests

---
title: "Rank Comparison"
author: "Blinded"
date: "2023-12-01"
output: pdf_document
---

The code below reads in the data comparing the 2022 National Competitions to the
2022 World Championships.
```{r}
national <- read.csv("../Data/ClusrankNatvsWorld.csv")
colnames(national) <- c("Time", "Athlete", "Stage")
```

# Overview
This analysis consists of twelve comparisons looking at all combinations of
the men's 100 meter dash, men's 110 meter hurdle, women's 100 meter dash, and
women's 100 meter hurdle.  In each comparison we will first read in the data
and then perform two clustered Wilcox tests, the first one without using
exact = TRUE, and the second one with exact = TRUE and 1,000,000 random
permutations. The tests compare the times and group by the Stage of the
competition, which in the data is coded as 1 if it took place during the 2022
World Championship and 0 otherwise. The tests cluster by athlete, of which for
each data set an integer from 1-11 is used (the values are arbitrary).
For more information about the "ds" method, please look at the Rank-based
Comparison subsection of the Methods section of the paper or
[at this link here.](https://www.jstatsoft.org/article/view/v096i06.)

First we will look compare the reaction times from athletes who competed in both
the 2019 and 2022 World Track and Field Championship.

# 2019 versus 2022

### Men's 110 meter hurdles

First we read in the data from the appropriate source.

```{r}
men_hurdle_2019 <- read.csv("../Data/Clusrank2019vs2022.csv")
colnames(men_hurdle_2019) <- c("Time", "Athlete", "Year")
```

Then we perform the clustered Wilcox tests.

```{r}
set.seed(1)
library(clusrank)
clusWilcox.test(Time, group=Year, cluster = Athlete, data = men_hurdle_2019,
                method = "ds")

clusWilcox.test(Time, group=Year, cluster = Athlete, data = men_hurdle_2019,
                method = "ds",
                exact = TRUE, B = 1000000)
```

### Women's 100 meter hurdles

First we read in the data from the appropriate source.

```{r}
women_hurdle_2019 <- read.csv("../Data/Clusrank2019vs2022Womens.csv")
colnames(women_hurdle_2019) <- c("Time", "Athlete", "Year")
```

Then we perform the clustered Wilcox tests.

```{r}
set.seed(1)
library(clusrank)
clusWilcox.test(Time, group=Year, cluster = Athlete, data = women_hurdle_2019,
                method = "ds")

clusWilcox.test(Time, group=Year, cluster = Athlete, data = women_hurdle_2019,
                method = "ds",
                exact = TRUE, B = 1000000)
```

### Men's 110 meter dash

First we read in the data from the appropriate source.

```{r}
men_dash_2019 <- read.csv("../Data/Clusrank2019vs2022Dash.csv")
colnames(men_dash_2019) <- c("Time", "Athlete", "Year")
```

Then we perform the clustered Wilcox tests.

```{r}
set.seed(1)
library(clusrank)
clusWilcox.test(Time, group=Year, cluster = Athlete, data = men_dash_2019,
                method = "ds")

clusWilcox.test(Time, group=Year, cluster = Athlete, data = men_dash_2019,
                method = "ds",
                exact = TRUE, B = 1000000)
```
### Women's 100 meter dash

First we read in the data from the appropriate source.

```{r}
women_dash_2019 <- read.csv("../Data/Clusrank2019vs2022WomensDash.csv")
colnames(women_dash_2019) <- c("Time", "Athlete", "Year")
```

Then we perform the clustered Wilcox tests.

```{r}
set.seed(1)
library(clusrank)
clusWilcox.test(Time, group=Year, cluster = Athlete, data = women_dash_2019,
                method = "ds")

clusWilcox.test(Time, group=Year, cluster = Athlete, data = women_dash_2019,
                method = "ds",
                exact = TRUE, B = 1000000)
```

# 2022 National versus World
Now we will look compare the reaction times from athletes at their 2022 national
track and field championships with the 2022 World Championships.

### Men's 110 meter hurdles

First we read in the data from the appropriate source.

```{r}
men_hurdle_2022 <- read.csv("../Data/ClusrankNatvsWorld.csv")
colnames(men_hurdle_2022) <- c("Time", "Athlete", "Year")
```

Then we perform the clustered Wilcox tests.

```{r}
set.seed(1)
library(clusrank)
clusWilcox.test(Time, group=Year, cluster = Athlete, data = men_hurdle_2022,
                method = "ds")

clusWilcox.test(Time, group=Year, cluster = Athlete, data = men_hurdle_2022,
                method = "ds",
                exact = TRUE, B = 1000000)
```

### Women's 100 meter hurdles

First we read in the data from the appropriate source.

```{r}
women_hurdle_2022 <- read.csv("../Data/ClusrankNatvsWorldWomens.csv")
colnames(women_hurdle_2022) <- c("Time", "Athlete", "Year")
```

Then we perform the clustered Wilcox tests.

```{r}
set.seed(1)
library(clusrank)
clusWilcox.test(Time, group=Year, cluster = Athlete, data = women_hurdle_2022,
                method = "ds")

clusWilcox.test(Time, group=Year, cluster = Athlete, data = women_hurdle_2022,
                method = "ds",
                exact = TRUE, B = 1000000)
```

### Men's 110 meter dash

First we read in the data from the appropriate source.

```{r}
men_dash_2022 <- read.csv("../Data/ClusrankNatvsWorldDash.csv")
colnames(men_dash_2022) <- c("Time", "Athlete", "Year")
```

Then we perform the clustered Wilcox tests.

```{r}
set.seed(1)
library(clusrank)
clusWilcox.test(Time, group=Year, cluster = Athlete, data = men_dash_2022,
                method = "ds")

clusWilcox.test(Time, group=Year, cluster = Athlete, data = men_dash_2022,
                method = "ds",
                exact = TRUE, B = 1000000)
```
### Women's 100 meter dash

First we read in the data from the appropriate source.

```{r}
women_dash_2022 <- read.csv("../Data/ClusrankNatvsWorldWomensDash.csv")
colnames(women_dash_2022) <- c("Time", "Athlete", "Year")
```

Then we perform the clustered Wilcox tests.

```{r}
set.seed(1)
library(clusrank)
clusWilcox.test(Time, group=Year, cluster = Athlete, data = women_dash_2022,
                method = "ds")

clusWilcox.test(Time, group=Year, cluster = Athlete, data = women_dash_2022,
                method = "ds",
                exact = TRUE, B = 1000000)
```
# 2022 versus 2023

Lastly we will look compare the reaction times from athletes who competed in both
the 2022 and 2023 World Track and Field Championship.

### Men's 110 meter hurdles

First we read in the data from the appropriate source.

```{r}
men_hurdle_2023 <- read.csv("../Data/Clusrank2022vs2023.csv")
colnames(men_hurdle_2023) <- c("Time", "Athlete", "Year")
```

Then we perform the clustered Wilcox tests.

```{r}
set.seed(1)
library(clusrank)
clusWilcox.test(Time, group=Year, cluster = Athlete, data = men_hurdle_2023,
                method = "ds")

clusWilcox.test(Time, group=Year, cluster = Athlete, data = men_hurdle_2023,
                method = "ds",
                exact = TRUE, B = 1000000)
```

### Women's 100 meter hurdles

First we read in the data from the appropriate source.

```{r}
women_hurdle_2023 <- read.csv("../Data/Clusrank2022vs2023Womens.csv")
colnames(women_hurdle_2023) <- c("Time", "Athlete", "Year")
```

Then we perform the clustered Wilcox tests.

```{r}
set.seed(1)
library(clusrank)
clusWilcox.test(Time, group=Year, cluster = Athlete, data = women_hurdle_2023,
                method = "ds")

clusWilcox.test(Time, group=Year, cluster = Athlete, data = women_hurdle_2023,
                method = "ds",
                exact = TRUE, B = 1000000)
```

### Men's 110 meter dash

First we read in the data from the appropriate source.

```{r}
men_dash_2023 <- read.csv("../Data/Clusrank2022vs2023Dash.csv")
colnames(men_dash_2023) <- c("Time", "Athlete", "Year")
```

Then we perform the clustered Wilcox tests.

```{r}
set.seed(1)
library(clusrank)
clusWilcox.test(Time, group=Year, cluster = Athlete, data = men_dash_2023,
                method = "ds")

clusWilcox.test(Time, group=Year, cluster = Athlete, data = men_dash_2023,
                method = "ds",
                exact = TRUE, B = 1000000)
```
### Women's 100 meter dash

First we read in the data from the appropriate source.

```{r}
women_dash_2023 <- read.csv("../Data/Clusrank2022vs2023WomensDash.csv")
colnames(women_dash_2023) <- c("Time", "Athlete", "Year")
```

Then we perform the clustered Wilcox tests.

```{r}
set.seed(1)
library(clusrank)
clusWilcox.test(Time, group=Year, cluster = Athlete, data = women_dash_2023,
                method = "ds")

clusWilcox.test(Time, group=Year, cluster = Athlete, data = women_dash_2023,
                method = "ds",
                exact = TRUE, B = 1000000)
```