### 2024-10-30 Code/NewAnalysis.Rmd
DQ analysis is at the bottom

---
title: "NewAnalysis"
author: "Owen Fiore"
date: "2024-07-03"
output: html_document
---

# New Work
To try and keep everything organized, we are going to use this document to test
anything additional before putting it in `GLMM.Rmd.`  We have several things
to test before our re-submission.

## Can we pool the dash and hurdles results?
We need to see how different our results are using a pooled version of the men's
100 meter dash and men's 110 meter hurdles.  We will also include a random effect
into our model to compensate for this.

```{r}
library(tidyverse)
times <- read.csv("../Data/GLMMData.csv")
colnames(times) <- c("Venue", "Type", "Time", "RxnTime", "Gender", "Heat", "Event")
```

```{r}
dash_and_hurdles <- times |>
  filter(Type == "F" | Type == "S") |>
  filter(Gender == "M") |>
  filter(Time != "DNS" & Time != "DQ" & Time != "D") |>
  filter(RxnTime != 0) |>
  filter(Event == "100 Dash" | Event == "110 Hurdles")

hurdles <- dash_and_hurdles |>
  filter(Event == "110 Hurdles")
```

We consider the baseline models:
```{r}
library(lme4)
dash_and_hurdles_one_gamma_log <- glm(RxnTime ~ 1, data = dash_and_hurdles, 
                                    family = Gamma(link = "log"))
hurdles_one_gamma_log <- glm(RxnTime ~ 1, data = hurdles, 
                                family = Gamma(link = "log"))
```
To assess whether we we can pool the data, we are going to perform an anova test
to determine whether the most recently added coefficient to the model was
significant.

```{r}
dash_and_hurdles_three_effect_gamma_log <- glmer(RxnTime ~ (1 | Venue) +
                                                   (1 | Heat) + Event, 
                                   data = dash_and_hurdles,
                                   family = Gamma(link = "log"))
dash_and_hurdles_two_effect_gamma_log <- glmer(RxnTime ~ (1 | Venue) +
                                                   (1 | Heat), 
                                   data = dash_and_hurdles,
                                   family = Gamma(link = "log"))

hurdles_two_effect_gamma_log <- glmer(RxnTime ~ (1 | Venue) + (1 | Heat), 
                               data = hurdles,
                               family = Gamma(link = "log"))
```

```{r}
anova(dash_and_hurdles_two_effect_gamma_log, dash_and_hurdles_three_effect_gamma_log)
```
What we observe is that there is not a significant difference between these
models as the p value is quite large. Thus we will not include `Event` as a
feature in our final model.

## 200m test
Does it make sense to include the 200m dash data as well?

```{r}
all_events <- times |>
  filter(Type == "F" | Type == "S") |>
  filter(Gender == "M") |>
  filter(Time != "DNS" & Time != "DQ") |>
  filter(RxnTime != 0) |>
  filter(Event == "100 Dash" | Event == "110 Hurdles" | Event == "200 Dash") |>
  mutate(RxnTime = ifelse(RxnTime < 0.1, 0.1, RxnTime))
```

```{r}
all_events_three_effect_gamma_log <- glmer(RxnTime ~ (1 | Venue) +
                                                   (1 | Heat) + Event, 
                                   data = all_events,
                                   family = Gamma(link = "log"))
all_events_two_effect_gamma_log <- glmer(RxnTime ~ (1 | Venue) +
                                                   (1 | Heat), 
                                   data = all_events,
                                   family = Gamma(link = "log"))
```

```{r}
anova(all_events_three_effect_gamma_log, all_events_two_effect_gamma_log)
```
Now we see that the event become significant

```{r}
library(car)
TukeyHSD(aov(RxnTime ~ Event, data = all_events))
```
Using a Tukey's Honestly Significant Difference test, we see that the 200m dash
differs significantly from 100m dash/hurdles.

Based on all the following evidence we are going to perform our men's analysis
on data from the 100 meter dash and 110 meter hurdles exclusively.

## Now for women
We want to make sure we get similar results when we pool the women's 100m hurdle
and 100m dash. 
```{r}
times <- read.csv("../Data/GLMMData.csv")
colnames(times) <- c("Venue", "Type", "Time", "RxnTime", "Gender", "Heat", "Event")
```

Again we are going to run an anova test to determine how impactful the coefficient
on `Event` is.
```{r}
women_dash_and_hurdles <- times |>
  filter(Type == "F" | Type == "S") |>
  filter(Gender == "F") |>
  filter(Time != "DNS") |>
  filter(RxnTime != 0) |>
  filter(Event == "100 Dash" | Event == "100 Hurdles") |>
  mutate(RxnTime = ifelse(RxnTime < 0.1, 0.1, RxnTime))
```

```{r}
women_dash_and_hurdles_three_effect_gamma_log <- glmer(RxnTime ~ (1 | Venue) +
                                                   (1 | Heat) + (1 | Event), 
                                   data = women_dash_and_hurdles,
                                   family = Gamma(link = "log"))
women_dash_and_hurdles_two_effect_gamma_log <- glmer(RxnTime ~ (1 | Venue) +
                                                   (1 | Heat), 
                                   data = women_dash_and_hurdles,
                                   family = Gamma(link = "log"))
```

```{r}
anova(women_dash_and_hurdles_two_effect_gamma_log, women_dash_and_hurdles_three_effect_gamma_log)
```
There seems to be more of a difference but the test is still not significant.

Based on the above evidence it seems reasonable to conclude that it is appropriate
to pool the men's 110 meter hurdles with the 100 meter dash and pool the women's
100 meter hurdles with the 100 dash but based on previous studies we are going
to separate men's and women's data.

## Testing Whether or not to include disqualfications
```{r}
dash_and_hurdles <- times |>
  filter(Type == "F" | Type == "S") |>
  filter(Gender == "M") |>
  filter(Time != "DNS" & Time != "DQ" & Time != "D") |>
  filter(RxnTime != 0) |>
  filter(Event == "100 Dash" | Event == "110 Hurdles")
```

```{r}
cat("The number of entries of the men's 100 meter dash and 110 meter hurdles
    without disqualifcations is", nrow(dash_and_hurdles))
```

```{r}
dash_and_hurdles_two_effect_gamma_log <- glmer(RxnTime ~ (1 | Venue) +
                                                   (1 | Heat), 
                                   data = dash_and_hurdles,
                                   family = Gamma(link = "log"))
```

We need the baseline model for our analysis as well
```{r}
dash_and_hurdles_one_gamma_log <- glm(RxnTime ~ 1, data = dash_and_hurdles, 
                                    family = Gamma(link = "log"))
```

```{r}
sim_two_effect_time <- function(model, base, RxnTime) {
  simfit <- function(model, n = 10000000) {
    vc <- as.data.frame(VarCorr(model))
    sd_Venue <- vc$sdcor[1]
    sd_Venue_heat <- vc$sdcor[2]

    z1 <- rnorm(n, sd = sd_Venue)
    z2 <- rnorm(n, sd = sd_Venue_heat)

    shape <- 1 / sigma(model)^2
    mu <- exp(fixef(model) + z1 + z2)

    x <- rgamma(n, shape = shape, scale = mu / shape)
    x
  }

  sbase <- summary(base)
  x <- simfit(model)
  print(mean(x < RxnTime))
  #print(quantile(x, prob = c(0.00001, 0.0001, 0.0005, 0.001)))
  ## compare with and without random effect
  #plot(density(x))  # with random effects
  #curve(dgamma(x, shape = 1 / sbase$dispersion, 
   #            scale = exp(coef(base)) * sbase$dispersion), 
  #      0.08, 0.25, add = TRUE, col = "blue")  # fixed effect only
}
```

```{r}
set.seed(1)
sim_two_effect_time(dash_and_hurdles_two_effect_gamma_log, dash_and_hurdles_one_gamma_log, 0.1)
```
Now lets see how our results change when we include disqualifications
```{r}
dash_and_hurdles_with_dq <- times |>
  filter(Type == "F" | Type == "S") |>
  filter(Gender == "M") |>
  filter(RxnTime != 0) |>
  filter(Event == "100 Dash" | Event == "110 Hurdles") |>
  mutate(RxnTime = ifelse(RxnTime < 0, 0.000001, RxnTime))
```

```{r}
cat("The number of entries of the men's 100 meter dash and 110 meter hurdles
    without disqualifcations is", nrow(dash_and_hurdles_with_dq))
```

```{r}
dash_and_hurdles_with_dq_two_effect_gamma_log <- glmer(RxnTime ~ (1 | Venue) +
                                                   (1 | Heat), 
                                   data = dash_and_hurdles_with_dq,
                                   family = Gamma(link = "log"))
```

We need the baseline model for our analysis as well
```{r}
dash_and_hurdles_with_dq_one_gamma_log <- glm(RxnTime ~ 1, data = dash_and_hurdles_with_dq, 
                                    family = Gamma(link = "log"))
```

```{r}
set.seed(1)
sim_two_effect_time(dash_and_hurdles_with_dq_two_effect_gamma_log, dash_and_hurdles_with_dq_one_gamma_log, 0.1)
```
## Lets repeat the previous section but exclude 2022 from our results
```{r}
dash_and_hurdles_no2022 <- times |>
  filter(Type == "F" | Type == "S") |>
  filter(Gender == "M") |>
  filter(Time != "DNS" & Time != "DQ" & Time != "D") |>
  filter(RxnTime != 0) |>
  filter(Event == "100 Dash" | Event == "110 Hurdles") |>
  filter(Venue != 2022)
```

```{r}
dash_and_hurdles_no2022_two_effect_gamma_log <- glmer(RxnTime ~ (1 | Venue) +
                                                   (1 | Heat), 
                                   data = dash_and_hurdles_no2022,
                                   family = Gamma(link = "log"))

dash_and_hurdles_no2022_one_gamma_log <- glm(RxnTime ~ 1, data = dash_and_hurdles_no2022, 
                                    family = Gamma(link = "log"))
```

```{r}
set.seed(1)
sim_two_effect_time(dash_and_hurdles_no2022_two_effect_gamma_log, dash_and_hurdles_no2022_one_gamma_log, 0.1)
```

Now we analyze without 2022 but with disqualifications
```{r}
dash_and_hurdles_no2022_with_dq <- times |>
  filter(Type == "F" | Type == "S") |>
  filter(Gender == "M") |>
  filter(RxnTime != 0) |>
  filter(Event == "100 Dash" | Event == "110 Hurdles") |>
  filter(Venue != 2022) |>
  mutate(RxnTime = ifelse(RxnTime < 0, 0.000001, RxnTime))
```

```{r}
dash_and_hurdles_no2022_with_dq_two_effect_gamma_log <- glmer(RxnTime ~ (1 | Venue) +
                                                   (1 | Heat), 
                                   data = dash_and_hurdles_no2022_with_dq,
                                   family = Gamma(link = "log"))

dash_and_hurdles_no2022_with_dq_one_gamma_log <- glm(RxnTime ~ 1, data = dash_and_hurdles_no2022_with_dq, 
                                    family = Gamma(link = "log"))
```

```{r}
sim_two_effect_time(dash_and_hurdles_no2022_with_dq_two_effect_gamma_log, dash_and_hurdles_no2022_with_dq_one_gamma_log, 0.1)
```