### 2024-11-20 Archive/Plots.Rmd
Overhauled document to only include relevant plots, listed in order of appearance currently in the manuscript

---
title: "Plots"
author: "Owen Fiore"
date: "2022-12-09"
output: html_document
---

```{r}
times <- read.csv("../Data/GLMMData.csv")
colnames(times) <- c("Venue", "Type", "Time", "RxnTime", "Gender", "Heat", "Event")
```

```{r}
library(tidyverse)
df <- times |>
  filter(Type == "F" | Type == "S") |>
  filter(Gender == "M") |>
  filter(Time != "DNS") |>
  filter(RxnTime != 0) |>
  filter(Event == "100 Dash" | Event == "110 Hurdles")
```

# Figure 1

### Mens 2019 vs 2022

```{r}
library(tidyr)
library(ggplot2)
nineteen <- read.csv("../Archive/2019vs2022.csv")
dash_19 <- read.csv("../Archive/2019vs2022Dash.csv")
#"C:\Users\ofior\Documents\Thesis\Data\2019vs2022.csv"
colnames(nineteen) <- c("Athlete", "Country", "2019 H", "2019 S", "2019 F", "2022 H", "2022 S", "2022 F")
colnames(dash_19) <- c("Athlete", "2019 H", "2019 S", "2019 F", "2022 H", "2022 S", "2022 F")
nineteen <- nineteen[, -c(1,2)]
dash_19 <- dash_19[, -c(1)]
nineteen <- rbind(nineteen, dash_19)

df_long <- tidyr::pivot_longer(nineteen, cols = everything(), names_to = "Column", values_to = "Time")
df_long$Column <- factor(df_long$Column, levels = c("2019 H", "2019 S", "2019 F", "2022 H", "2022 S", "2022 F"))
men_19 <- ggplot(df_long, aes(x = Column, y = Time)) +
  geom_point() +
  labs(y = "Reaction Time", x = "") +
  annotate("text", y = 0.21, x = "2022 S", label = "Men 2019 vs 2022", cex = 3) +
  ylim(0.08, 0.22) +
  theme_minimal()
```

### Womens 2019 vs 2022

```{r}
library(tidyr)
library(ggplot2)
nineteen_w <- read.csv("../Archive/2019vs2022Womens.csv")
dash_19_w <- read.csv("../Archive/2019vs2022WomensDash.csv")
#"C:\Users\ofior\Documents\Thesis\Data\2019vs2022.csv"
colnames(nineteen_w) <- c("Athlete", "Country", "2019 H", "2019 S", "2019 F", "2022 H", "2022 S", "2022 F")
colnames(dash_19_w) <- c("Athlete", "2019 H", "2019 S", "2019 F", "2022 H", "2022 S", "2022 F")
nineteen_w <- nineteen_w[, -c(1,2, 9)]
dash_19_w <- dash_19_w[, -c(1)]
nineteen_w <- rbind(nineteen_w, dash_19_w)

df_long <- tidyr::pivot_longer(nineteen_w, cols = everything(), names_to = "Column", values_to = "Time")
df_long$Column <- factor(df_long$Column, levels = c("2019 H", "2019 S", "2019 F", "2022 H", "2022 S", "2022 F"))
women_19 <- ggplot(df_long, aes(x = Column, y = Time)) +
  geom_point() +
  labs(y = "Reaction Time", x = "") +
  annotate("text", y = 0.21, x = "2022 S", label = "Women 2019 vs 2022", cex = 3) +
  ylim(0.08, 0.22) +
  theme_minimal()
```

### 2022 Mens National vs International

```{r}
library(tidyr)
library(ggplot2)
df1 <- read.csv("../Archive/NatVSWorld.csv")
df2 <- read.csv("../Archive/2022NatvsWorldDash.csv")
#"C:\Users\ofior\Documents\Thesis\Data\2019vs2022.csv"
colnames(df1) <- c("Athlete", "Country", "Nat. H", "Nat. S", "Nat. F", "Inter. H", "Inter. S", "Inter. F")
colnames(df2) <- c("Athlete", "Country", "Nat. H", "Nat. S", "Nat. F", "Inter. H", "Inter. S", "Inter. F")
df1 <- df1[, -c(1,2, 9)]
df2 <- df2[, -c(1, 2)]
df <- rbind(df1, df2)

df_long <- tidyr::pivot_longer(df, cols = everything(), names_to = "Column", values_to = "Time")
df_long$Column <- factor(df_long$Column, levels = c("Nat. H", "Nat. S", "Nat. F", "Inter. H", "Inter. S", "Inter. F"))
men_22 <- ggplot(df_long, aes(x = Column, y = Time)) +
  geom_point() +
  labs(y = "Reaction Time", x = "") +
  annotate("text", y = 0.21, x = "Inter. S", label = "Men 2022 Nat. vs Inter.", cex = 3) +
  ylim(0.08, 0.22) +
  theme_minimal()
```

### Womens 2022 National vs International

```{r}
library(tidyr)
library(ggplot2)
df1 <- read.csv("../Archive/2022NatvsWorldWomens.csv")
df2 <- read.csv("../Archive/2022NatvsWorldWomensDash.csv")
#"C:\Users\ofior\Documents\Thesis\Data\2019vs2022.csv"
colnames(df1) <- c("Athlete", "Country", "Nat. H", "Nat. S", "Nat. F", "Inter. H", "Inter. S", "Inter. F")
colnames(df2) <- c("Athlete", "Country", "Nat. H", "Nat. S", "Nat. F", "Inter. H", "Inter. S", "Inter. F")
df1 <- df1[, -c(1, 2)]
df2 <- df2[, -c(1, 2)]
df <- rbind(df1, df2)

df_long <- tidyr::pivot_longer(df, cols = everything(), names_to = "Column", values_to = "Time")
df_long$Column <- factor(df_long$Column, levels = c("Nat. H", "Nat. S", "Nat. F", "Inter. H", "Inter. S", "Inter. F"))
women_22 <- ggplot(df_long, aes(x = Column, y = Time)) +
  geom_point() +
  labs(y = "Reaction Time", x = "") +
  annotate("text", y = 0.21, x = "Inter. S", label = "Women 2022 Nat. vs Inter.", cex = 3) +
  ylim(0.08, 0.22) +
  theme_minimal()
```

### Men's 2022 vs 2023

```{r}
library(tidyr)
library(ggplot2)
df1 <- read.csv("../Archive/2022vs2023.csv")
df2 <- read.csv("../Archive/2022vs2023Dash.csv")
#"C:\Users\ofior\Documents\Thesis\Data\2019vs2022.csv"
colnames(df1) <- c("Athlete", "Country", "2022 H", "2022 S", "2022 F", "2023 H", "2023 S", "2023 F")
colnames(df2) <- c("Athlete", "2022 H", "2022 S", "2022 F", "2023 H", "2023 S", "2023 F")
df1 <- df1[, -c(1, 2)]
df2 <- df2[, -c(1)]
df <- rbind(df1, df2)

df_long <- tidyr::pivot_longer(df, cols = everything(), names_to = "Column", values_to = "Time")
df_long$Column <- factor(df_long$Column, levels = c("2023 H", "2023 S", "2023 F", "2022 H", "2022 S", "2022 F"))
men_23 <- ggplot(df_long, aes(x = Column, y = Time)) +
  geom_point() +
  labs(y = "Reaction Time", x = "") +
  annotate("text", y = 0.21, x = "2022 S", label = "Men 2022 vs 2023", cex = 3) +
  ylim(0.08, 0.22) +
  theme_minimal()
```

### Womens 2022 vs 2023

```{r}
library(tidyr)
library(ggplot2)
df1 <- read.csv("../Archive/2022vs2023Womens.csv")
df2 <- read.csv("../Archive/2022vs2023WomensDash.csv")
#"C:\Users\ofior\Documents\Thesis\Data\2019vs2022.csv"
colnames(df1) <- c("Athlete", "Country", "2022 H", "2022 S", "2022 F", "2023 H", "2023 S", "2023 F")
colnames(df2) <- c("Athlete", "2022 H", "2022 S", "2022 F", "2023 H", "2023 S", "2023 F")
df1 <- df1[, -c(1, 2)]
df2 <- df2[, -c(1)]
df <- rbind(df1, df2)

df_long <- tidyr::pivot_longer(df, cols = everything(), names_to = "Column", values_to = "Time")
df_long$Column <- factor(df_long$Column, levels = c("2023 H", "2023 S", "2023 F", "2022 H", "2022 S", "2022 F"))
women_23 <- ggplot(df_long, aes(x = Column, y = Time)) +
  geom_point() +
  labs(y = "Reaction Time", x = "") +
  annotate("text", y = 0.21, x = "2022 S", label = "Women 2022 vs 2023", cex = 3) +
  ylim(0.08, 0.22) +
  theme_minimal()
```

### Combines the previous 6 graphs onto 1 plot

```{r}
library(gridExtra)
rank_scatter <- grid.arrange(men_19, women_19, men_22,women_22, men_23, women_23, ncol = 2)
ggsave(filename = "RankScatterPlots.pdf", path = "../Manuscript", plot=rank_scatter, height = 7.5, width = 7)
```

# Figure 2

```{r}
library(ggplot2)
ggplot(df, aes(x = Venue, y = RxnTime, group = Venue)) +
  geom_boxplot() +
  labs(x = "Year", y = "Reaction Time (s)") +
       scale_y_continuous(limits = c(0.05, 0.275),
                            breaks = seq(0.05, 0.275, 0.025)) +
  expand_limits(y = 0.1) + scale_x_continuous(limits = c(1998,2024),
                                              breaks = seq(1999, 2024, 2))
ggsave(filename="Boxplot.pdf", path = "../Manuscript", height = 4, width = 6)
```

# Figure 3

setup
```{r}
library("gamlss")
library("qqconf")
dhm <- times |>
  filter(Type == "F" | Type == "S") |>
  filter(Gender == "M") |>
  filter(Time != "DNS") |>
  filter(RxnTime > 0) |>
  filter(Event == "100 Dash" | Event == "110 Hurdles")

dhm$Venue <- as.factor(dhm$Venue)
dhm$Heat <- as.factor(dhm$Heat)
gg3b  <- gamlss(RxnTime ~ random(Venue), sigma.formula = ~ random(Heat), data = dhm, family = GG, control = gamlss.control(n.cyc = 40))

dhm_no2022 <- times |>
  filter(Type == "F" | Type == "S") |>
  filter(Gender == "M") |>
  filter(Time != "DNS") |>
  filter(RxnTime > 0) |>
  filter(Event == "100 Dash" | Event == "110 Hurdles") |>
  filter(Venue != 2022) |>
  mutate(Venue = as.factor(Venue), Heat = as.factor(Heat))

gg3b_no2022  <- gamlss(RxnTime ~ random(Venue), sigma.formula = ~ random(Heat), data = dhm_no2022, family = GG, control = gamlss.control(n.cyc = 40))

```

plot 1
```{r}
venue_effects <- as.data.frame(gg3b$mu.coefSmo[[1]]$coef)
colnames(venue_effects) <- c("Effect Value")

venue_plot <- ggplot(venue_effects, aes(x = rownames(venue_effects),
                                               y = `Effect Value`)) +
  geom_point() + 
  labs(x = "Venue", y = "Venue Effect")+
  scale_y_continuous(limits = c(-0.15, 0.1))
```
plot 2
```{r}
venue_effects_no2022 <- as.data.frame(gg3b_no2022$mu.coefSmo[[1]]$coef)
venue_effects_no2022["2022", ] <- 1 #Arbitrary number out of range of y scale so
#that the x axis line up

colnames(venue_effects_no2022) <- c("Effect Value")

venue_plot_no2022 <- ggplot(venue_effects_no2022,
                                   aes(x = rownames(venue_effects_no2022),
                                       y = `Effect Value`)) +
  geom_point() + 
  labs(x = "Venue", y = "Venue Effect") +
  scale_y_continuous(limits = c(-0.15, 0.1))
```
### Now we combine for final product
```{r}
library(gridExtra)
venue_effects_combined <- grid.arrange(venue_plot,
                                       venue_plot_no2022, ncol = 1)
ggsave(filename = "ComparisonOfVenueEffects.pdf", path = "../Manuscript",
       plot=venue_effects_combined, height = 6.5, width = 7)
```
# Figure 4
```{r}

pdf(file = "../Manuscript/GGResiduals.pdf", height = 4, width = 6)
qq_conf_plot(residuals(gg3b), dparams = list(mean = 0, sd = 1))
dev.off()
```

# Figure 5
```{r}
sim_time <- function(model, RxnTime) {
  simfit <- function(model, n = 10000000) {

    sd_Venue <- model$mu.coefSmo[[1]]$sigb
    sd_Venue_heat <- model$sigma.coefSmo[[1]]$sigb[1]

    mu <- exp(model$mu.coefficients[1] + rnorm(n, mean = 0 , sd = sd_Venue))
    sigma <- exp(model$sigma.coefficients[1] + rnorm(n, mean = 0 , sd = sd_Venue_heat))
    nu <- model$nu.coefficients[1]

    x <- rGG(n, mu = mu, sigma = sigma, nu = nu)
    x
  }
  x <- simfit(model)
  plot(density(x), xlim = c(0, 0.3), main = "", xlab = "")
  curve(dGG(x, mu = exp(model$mu.coefficients[1]),
            sigma = exp(model$sigma.coefficients[1]),
            nu = model$nu.coefficients[1]),0, 0.3, add = TRUE, col = "blue",
        main = "", xlab = "")
}
pdf(file = "../Manuscript/GGDensity.pdf", height = 4, width = 6)
sim_time(gg3b, 0.1)
dev.off()
```