### 2024-11-29 Code/RankComparison.Rmd
Updated language and removed supplementary code

---
title: "Rank Comparison"
author: "Blinded"
date: "2023-12-01"
output: pdf_document
---

# Overview
This analysis consists of six comparisons of times comparing athletes who
competed at the 2022 World Championships and another major event. The six
analyses are listed below. Men's data contains values from athletes who competed
in either the men's 110 meter hurdles or 100 meter dash at championship level
events. Women's data contains values from athletes who competed
in either the women's 100 meter hurdles or 100 meter dash at championship level
events.

+ 2019 and 2022 World Championships men's data
+ 2019 and 2022 World Championships women's data
+ 2022 national championships and 2022 World Championships men's data
+ 2022 national championships and 2022 World Championships women's data
+ 2023 and 2022 World Championships men's data
+ 2023 and 2022 World Championships women's data

In each comparison we will first read in the data
and then perform two clustered Wilcox tests, the first one without using
exact = TRUE, and the second one with exact = TRUE and 1,000,000 random
permutations. The tests compare the times and group by the Stage of the
competition, and cluster by athlete.  An explanation of the features of each
data is below.

+ `Time` is the athlete's reaction time
+ `Athlete` is an arbitrary integer assigned to each athlete.  Each value is
unique within each data set.
+ `Year` is coded as 1 if it took place during the 2022 World Championship and
0 otherwise.
+ `Event` is coded as 0 for a hurdles event and 1 as a dash event, although this
has no impact on the analysis.

For more information about the "ds" method, please look at the Rank-based
Comparison subsection of the Methods section of the paper or
[at this link here.](https://www.tandfonline.com/doi/abs/10.1198/016214504000001583)

# 2019 versus 2022

First we will look compare the reaction times from athletes who competed in both
the 2019 and 2022 World Track and Field Championship.

### Men's Data

First we read in the data from the appropriate source.

```{r}
men_2019 <- read.csv("../Data/Clusrank2019vs2022.csv")
colnames(men_2019) <- c("Time", "Athlete", "Year", "Event")
```

Then we perform the clustered Wilcox tests.

```{r}
set.seed(1)
library(clusrank)
clusWilcox.test(Time, group=Year, cluster = Athlete, data = men_2019,
                method = "ds")

clusWilcox.test(Time, group=Year, cluster = Athlete, data = men_2019,
                method = "ds",
                exact = TRUE, B = 1000000)
```
The p-value from the asymptomatic test was $1.126 \cdot 10^{-5}$ and the p-value
from the permutation test was $2.8 \cdot 10 ^ {-5}$. Both of these are
statistically significant results showing a difference in male athletes who
competed in 2019 and 2022.

### Women's Data

First we read in the data from the appropriate source.

```{r}
women_2019 <- read.csv("../Data/Clusrank2019vs2022Womens.csv")
colnames(women_2019) <- c("Time", "Athlete", "Year", "Event")
```

Then we perform the clustered Wilcox tests.

```{r}
set.seed(1)
library(clusrank)
clusWilcox.test(Time, group=Year, cluster = Athlete, data = women_2019,
                method = "ds")

clusWilcox.test(Time, group=Year, cluster = Athlete, data = women_2019,
                method = "ds",
                exact = TRUE, B = 1000000)
```
The p-value from the asymptomatic test was $6.9 \cdot 10^{-3}$ and the p-value
from the permutation test was $1.5 \cdot 10^{-3}$. Both of these are
statistically significant results showing a difference in female athletes who
competed in 2019 and 2022.

# 2022 National versus World
Now we will look compare the reaction times from athletes at their 2022 national
track and field championships with the 2022 World Championships.

### Men's Data

First we read in the data from the appropriate source.

```{r}
men_2022 <- read.csv("../Data/ClusrankNatvsWorld.csv")
colnames(men_2022) <- c("Time", "Athlete", "Year", "Event")
```

Then we perform the clustered Wilcox tests.

```{r}
set.seed(1)
library(clusrank)
clusWilcox.test(Time, group=Year, cluster = Athlete, data = men_2022,
                method = "ds")

clusWilcox.test(Time, group=Year, cluster = Athlete, data = men_2022,
                method = "ds",
                exact = TRUE, B = 1000000)
```

The p-value from the asymptomatic test was $6.1 \cdot 10^{-5}$ and the p-value
from the permutation test was $1.0 \cdot 10^{-6}$. Both of these are
statistically significant results showing a difference in male athletes who
competed at various competitions in 2022.

### Women's Data

First we read in the data from the appropriate source.

```{r}
women_2022 <- read.csv("../Data/ClusrankNatvsWorldWomens.csv")
colnames(women_2022) <- c("Time", "Athlete", "Year", "Event")
```

Then we perform the clustered Wilcox tests.

```{r}
set.seed(1)
library(clusrank)
clusWilcox.test(Time, group=Year, cluster = Athlete, data = women_2022,
                method = "ds")

clusWilcox.test(Time, group=Year, cluster = Athlete, data = women_2022,
                method = "ds",
                exact = TRUE, B = 1000000)
```
The p-value from the asymptomatic test was $1.2 \cdot 10^{-3}$ and the p-value
from the permutation test was $1.0 \cdot 10^{-6}$. Both of these are
statistically significant results showing a difference in female athletes who
competed at various competitions in 2022.

# 2022 versus 2023

Lastly we will look compare the reaction times from athletes who competed in both
the 2022 and 2023 World Track and Field Championship.

### Men's Data

First we read in the data from the appropriate source.

```{r}
men_2023 <- read.csv("../Data/Clusrank2022vs2023.csv")
colnames(men_2023) <- c("Time", "Athlete", "Year", "Event")
```

Then we perform the clustered Wilcox tests.

```{r}
set.seed(1)
library(clusrank)
clusWilcox.test(Time, group=Year, cluster = Athlete, data = men_2023,
                method = "ds")

clusWilcox.test(Time, group=Year, cluster = Athlete, data = men_2023,
                method = "ds",
                exact = TRUE, B = 1000000)
```

The p-value from the asymptomatic test was $1.4 \cdot 10^{-6}$ and the p-value
from the permutation test was $2.8 \cdot 10 ^ {-5}$. Both of these are
statistically significant results showing a difference in male athletes who
competed in 2022 and 2023.

### Women's Data

First we read in the data from the appropriate source.

```{r}
women_2023 <- read.csv("../Data/Clusrank2022vs2023Womens.csv")
colnames(women_2023) <- c("Time", "Athlete", "Year", "Event")
```

Then we perform the clustered Wilcox tests.

```{r}
set.seed(1)
library(clusrank)
clusWilcox.test(Time, group=Year, cluster = Athlete, data = women_2023,
                method = "ds")

clusWilcox.test(Time, group=Year, cluster = Athlete, data = women_2023,
                method = "ds",
                exact = TRUE, B = 1000000)
```

The p-value from the asymptomatic test was $9.4 \cdot 10^{-7}$ and the p-value
from the permutation test was $1.5 \cdot 10^{-3}$. Both of these are
statistically significant results showing a difference in female athletes who
competed in 2019 and 2022.